Maximum Entropy Spectral Analysis: a case study

### 2021-03-24 paper/mesa.tex
Change in tex

@@

To show the analogy, we compute the PSD $S_{AR(p)}$ of an $AR(p)$ process and we show that it is formally equivalent to the PSD obtained in eq.~\eqref{eq:MESApsd}. This will also provide a direct expression for the autoregressive coefficients $b_i$ and for the noise variance $\sigma^2$.
We start taking the $z$ transform~\footnote{We remind the reader that the $z$ transform is the discrete-time equivalent of the Laplace transform, thus taking a discrete time time-series and returning a frequency series.} of Eq.~(\eqref{eq:AR_p}): 
\begin{align}
    \sum_t x_t z^t - \sum_i b_i z^i\sum_t x_{t - i} z^{t - i} = \sum_t \nu_t z^t.
\end{align}
Calling $\tilde x(z)$ and $\tilde \nu (z)$, the transformed quantities, 
in the $z$ domain, the process takes the form:
\begin{equation}
    \tilde x(z) = \frac{\tilde\nu(z)}{\left(1 - \sum_{i = n}^p b_n z^n \right)}
\end{equation}
Since we assumed a wide-sense stationary process, $\tilde{x}(z)$ is analytic both on and inside the unit circle. Taking its square value and evaluating it on the unit circle $z = e^{-\imath 2 \pi f \Delta t}$, from the definition of spectral density one obtains:
\begin{equation}\label{eq:ARspectrum}
    S_{AR(p)}(f) = \vert \tilde x(z)\vert ^ 2 = 
    \frac{\vert \tilde \nu(f) \vert ^ 2}{\left\vert 1 - \sum_{n = 1}^p b_i e^{\imath 2 \pi f n \Delta t} \right\vert ^ 2}\,.
\end{equation}
The numerator is the spectral density of white noise $\nu_t$, i.e. its (constant) variance $\sigma^2$.
Eq.~(\eqref{eq:ARspectrum}) and Eq.~(\eqref{eq:MESApsd}) are equivalent, if we identify $b_i = - a_i$ and $P_N \Delta t= \sigma ^ 2$.
This shows that the MAXENT estimation of the PSD also computes the coefficients of the autoregressive 
process that model the observed process.

@@

\pagebreak
\appendix
\section{Computing the PSD} 
\subsection{MESA solution}\label{sec:MESA_solution}
We derive the expression for the MAXENT spectral estimator following the approach proposed by Burg \cite{burg1975maximum}.
Unlike the standard approach, we do not enforce the constraints in eq.~\eqref{eq:MaxConstraint} with the standard Lagrange Multipliers approach.
We write instead the PSD $S(f)$ as the Fourier Transform of the sample autocorrelation function: 
\begin{equation}
    S(f) = \frac{1}{2 Ny}\sum_{n = -\infty}^{\infty} \bar r_n e^{- \imath 2 \pi n \Delta t},
\end{equation}
and the entropy gain expression eq.~\eqref{eq:EntropyGain} becomes:
\begin{equation}
    \Delta H = \int_{-Ny}^{Ny}  
    \log\left(\frac{1}{2 Ny}\sum_{n = -\infty}^{\infty} \bar r_n e^{-\imath 2 \pi f n \Delta t} 
    \right) df.
\end{equation}
Note that this expression already takes into account the contraints in eq.~\eqref{eq:MaxConstraint}.

>>>>

\sschmidt{Spiega bene i due metodi di windowing!}

@@

To show the analogy, we compute the PSD $S_{AR(p)}$ of an $AR(p)$ process and we show that it is formally equivalent to the PSD obtained in eq.~\eqref{eq:MESApsd}. This will also provide a direct expression for the autoregressive coefficients $b_i$ and for the noise variance $\sigma^2$.
We start taking the $z$ transform~\footnote{We remind the reader that the $z$ transform is the discrete-time equivalent of the Laplace transform, thus taking a discrete time time-series and returning a frequency series.} of Eq.~\eqref{eq:AR_p}: 
\begin{align}
    \sum_t x_t z^t - \sum_i b_i z^i\sum_t x_{t - i} z^{t - i} = \sum_t \nu_t z^t.
\end{align}
Calling $\tilde x(z)$ and $\tilde \nu (z)$, the transformed quantities, 
in the $z$ domain, the process takes the form:
\begin{equation}
    \tilde x(z) = \frac{\tilde\nu(z)}{\left(1 - \sum_{i = n}^p b_n z^n \right)}
\end{equation}
Since we assumed a wide-sense stationary process, $\tilde{x}(z)$ is analytic both on and inside the unit circle. Taking its square value and evaluating it on the unit circle $z = e^{-\imath 2 \pi f \Delta t}$, from the definition of spectral density one obtains:
\begin{equation}\label{eq:ARspectrum}
    S_{AR(p)}(f) = \vert \tilde x(z)\vert ^ 2 = 
    \frac{\vert \tilde \nu(f) \vert ^ 2}{\left\vert 1 - \sum_{n = 1}^p b_i e^{\imath 2 \pi f n \Delta t} \right\vert ^ 2}\,.
\end{equation}
The numerator is the spectral density of white noise $\nu_t$, i.e. its (constant) variance $\sigma^2$.
Eq.~(\eqref{eq:ARspectrum}) and Eq.~(\eqref{eq:MESApsd}) are equivalent, if we identify $b_i = - a_i$ and $P_N \Delta t= \sigma ^ 2$.
This shows that the MAXENT estimation of the PSD also computes the coefficients of the autoregressive 
process that model the observed process.

@@

\pagebreak
\appendix
\section{Details of PSD computation} 
\subsection{MESA solution}\label{sec:MESA_solution}
We derive the expression for the MAXENT spectral estimator following the approach proposed by Burg \cite{burg1975maximum}.
Unlike the standard approach, we do not enforce the constraints in eq.~\eqref{eq:MaxConstraint} with the standard Lagrange Multipliers approach.
We write instead the PSD $S(f)$ as the Fourier Transform of the sample autocorrelation function: 
\begin{equation}
    S(f) = \frac{1}{2 Ny}\sum_{n = -\infty}^{\infty} \bar r_n e^{- \imath 2 \pi n \Delta t},
\end{equation}
and the entropy gain expression eq.~\eqref{eq:EntropyGain} becomes:
\begin{equation}
    \Delta H = \int_{-Ny}^{Ny}  
    \log\left(\frac{1}{2 Ny}\sum_{n = -\infty}^{\infty} \bar r_n e^{-\imath 2 \pi f n \Delta t} 
    \right) df.
\end{equation}
Note that this expression already takes into account the contraints in eq.~\eqref{eq:MaxConstraint}.
